Supplementary Report for the article *Functional diversity can facilitate the collapse of an undesirable ecosystem state*

### 2023-01-26 reports/supplement/supplement_0.9.1.Rmd
various updates

```{r}
microxanox_version_here <- "0.9.1"
microxanox_version_path <- paste0("microxanox_v", microxanox_version_here)
if (packageVersion("microxanox") < package_version(microxanox_version_here)) {
  stop(paste0("microxanox version needs to be at least ", microxanox_version_here))
}
print(paste0("Version of microxanox: ", microxanox_version_here))
```

@@

## Investigating the mechanisms operating in our system

@@

          (num_strains == -1 & var_treat == "CB" & stand_var > 1.00) ~ "non-standard",
          (num_strains == -1 & var_treat == "SB" & stand_var > 0.60) ~ "non-standard",
          (num_strains == -1 & var_treat == "PB" & stand_var > 1.00) ~ "non-standard",
          (num_strains == -1 & var_treat == "CB-SB" & stand_var > 0.60) ~ "non-standard",
          (num_strains == -1 & var_treat == "CB-PB" & stand_var > 1.00) ~ "non-standard",
          (num_strains == -1 & var_treat == "SB-PB" & stand_var > 1.00) ~ "non-standard",
          (num_strains == -1 & var_treat == "CB-SB-PB" & stand_var > 1.00) ~ "non-standard"
         ),
         stab_states_type = ifelse(is.na(stab_states_type), "standard", "non-standard")
  )

@@

>>>>

```{r}
microxanox_version_here <- "0.9.1"
microxanox_version_path <- paste0("microxanox_v", microxanox_version_here)
if (packageVersion("microxanox") < package_version(microxanox_version_here)) {
  stop(paste0("microxanox version needs to be at least ", microxanox_version_here))
}
print(paste0("Version of microxanox: ", microxanox_version_here))

@@

## Investigating the mechanisms operating

@@

          (num_strains == -1 & var_treat == "CB" & stand_var > 1.00) ~ "non-standard",
          (num_strains == -1 & var_treat == "SB" & stand_var > 0.60) ~ "non-standard",
          (num_strains == -1 & var_treat == "PB" & stand_var > 1.00) ~ "non-standard",
          (num_strains == -1 & var_treat == "CB-SB" & stand_var > 0.60) ~ "non-standard",
          (num_strains == -1 & var_treat == "CB-PB" & stand_var > 1.00) ~ "non-standard",
          (num_strains == -1 & var_treat == "SB-PB" & stand_var > 1.00) ~ "non-standard",
          (num_strains == -1 & var_treat == "CB-SB-PB" & stand_var > 1.00) ~ "non-standard"
         ),
         stab_states_type = ifelse(is.na(stab_states_type), "standard", "non-standard")
  ) %>%
  mutate(hyst_min_log = ifelse(var_treat == "CB-SB" &
                                 stab_states_type == "non-standard" &
                                 Species == "SB_tot" &
                                 stand_var > 0.85,
                               -8,
                               hyst_min_log
                                 ))

@@

# Log and non-log oxygen diffusivity results

(ref:fig-log-nonlog) The effect of trait diversity (y-axis) viewed with a log axis (left panels) or arithmetic axis for the oxygen diffusivity gradient (x-axis). The positions of the tipping points (anoxic to oxic in blue, oxic to anoxic in red) and the extent of the region of bistability (green or orange shaded region) are shown. The left-most region is anoxic only, the middle is oxic or anoxic depending on historical conditions, and the right-most is oxic only. In the left panels the thick grey vertical lines at -8 and 0 on the x-axis show the extent of the oxygen diffusivity gradient investigated in the simulations. In panels e, f, i, and j the orange region shows when patterns of coexistence were atypical of the classical pattern shown in Figure 1a of the main text (e.g. when cyanobacteria coexisted with sulfur bacteria at low oxygen diffusivities). In the right-hand panels (arithmetic x-axis scale) the blue anoxic to oxic tipping point line often has steps. This is becase the simulated oxygen diffusivity levels were evenly distributed on the log-oxygen diffusivity scale, and so are relatively far apart at high oxygen diffusivities on the arithmetic scale.

```{r fig-log-nonlog, echo = FALSE,  include=TRUE, fig.height=8, fig.width=8, fig.align="center", fig.cap='(ref:fig-log-nonlog)'}
all_stab <- readRDS(here("data",
                         microxanox_version_path,
                         "temporal_method",
                         "event_definition_2",
                         "processed_data",
                         "all_stab_data.RDS"))

wait_time <- 1e6
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time)

plot_here <- plot_here %>%
  mutate(stab_states_type = "standard",
         stab_states_type = case_when(
           (var_treat == "CB" & stand_var > 1.00) ~ "non-standard",
           (var_treat == "SB" & stand_var > 1.00) ~ "non-standard",
           (var_treat == "PB" & stand_var > 0.65) ~ "non-standard",
           (var_treat == "CB-SB" & stand_var > 1.00) ~ "non-standard",
           (var_treat == "CB-PB" & stand_var > 0.65) ~ "non-standard",
           (var_treat == "SB-PB" & stand_var > 1.00) ~ "non-standard",
           (var_treat == "CB-SB-PB" & stand_var > 1.00) ~ "non-standard"
         ),
         stab_states_type = ifelse(is.na(stab_states_type), "standard", "non-standard")
  )

p1 <- fig_div_vs_o2diff_1strain_7row_stable_state_type(plot_here,
                                                       which_strain = num_strains,
                                                       figure_title = NULL
                                                       # figure_title = paste0(wait_time,
                                                       #                       " wait time,\n",
                                                       #                       num_strains,
                                                       #                       " strains")
)

p2 <- fig_div_vs_o2diff_1strain_7row_stable_state_type_nonlog(plot_here,
                                                       which_strain = num_strains,
                                                       figure_title = NULL
                                                       # figure_title = paste0(wait_time,
                                                       #                       " wait time,\n",
                                                       #                       num_strains,
                                                       #                       " strains")
)

p1 + p2
